---
title: "DataKind - BLR - Pratham Books Sprint - Exploratory Analysis"
author: "Arpit Arora"
output:
  html_document:
    highlight: tango
    theme: yeti
    toc: yes
    toc_depth: 6
---

#### Objective

1. Clean and normalize the data for analysis
2. Exploratory analysis of the stories data

#### Setting up the working environment

```{r setup}
# Setting up the working directory
knitr::opts_knit$set(root.dir = "~/Documents/DataKind/PrathamBooks-Sprint-2018/data")

# Load the required libraries
library(tidyverse)
library(formattable)

# Set plot theme
theme_set(theme_classic())
```

#### Reading the datasets

```{r read, message = FALSE, cache = TRUE, warning = FALSE}
# Reading the data files
stories <- read_csv("raw/stories.csv")

categories <- read_csv("raw/categories.csv")

tags <- read_csv("raw/tags.csv")

authors <- read_csv("raw/authors.csv")

illustrators <- read_csv("raw/illustrators.csv")

stories_pages_illustrations_master <- read_csv("raw/stories_pages_illustrations_master.csv")
```

```{r}
glimpse(stories_pages_illustrations_master)
```


#### Exploratory Data Analysis - Stories

##### Cleaning the dataset

```{r stories}
# Stories dataset
# Clean-up and transform dataset for analysis
stories <- stories %>%
                mutate(reading_level_updated = paste0("L", reading_level + 1),
                       # Normalize the reading_level field
                       ancestry_updated = ifelse(is.na(ancestry), story_id, ancestry),
                       # Normalize the ancestry field
                       derivation_type_updated = ifelse(is.na(derivation_type), "original", derivation_type),
                       # Get recommended story indicator
                       is_recommended_story = ifelse(is.na(publisher), F, T)
                      )

# Roll-up the categories dataset
categories_grouped <- aggregate(category_name ~ story_id, data = categories, paste, collapse = ", ")

# Roll-up the tags dataset
tags_grouped<- aggregate(tag_name ~ story_id, data = tags, paste, collapse = ", ")

# Clean up author dataset
authors <- authors %>%
                mutate(author_name = paste(author_first_name, author_last_name, sep = " ")) %>%
                select(story_id, author_name)

authors_grouped <- aggregate(author_name ~ story_id, data = authors, paste, collapse = ", ")

# Clean the illustrators dataset
illustrators <- illustrators %>%
                    mutate(illustrator_name = paste(illustrator_first_name, illustrator_last_name, sep = " ")) %>%
                    select(story_id, illustrator_name)
                
illustrators_grouped <- aggregate(illustrator_name ~ story_id, data = illustrators, paste, collapse = ", ")

# Clean ancestry for double parent stories
stories <- stories %>% 
                separate(ancestry_updated, c("ancestry_updated"), "/", extra = "drop", fill = "right")

stories$ancestry_updated <- as.numeric(stories$ancestry_updated)

# Processing data to creater master file for dashboard
# Join all the relevant data files
stories_master <- stories %>%
                    left_join(categories_grouped, by = "story_id") %>%
                    left_join(tags_grouped, by = "story_id") %>%
                    left_join(authors_grouped, by = "story_id") %>%
                    left_join(illustrators_grouped, by = "story_id") %>%
                    left_join(authors_grouped, by = c("ancestry_updated" = "story_id"), suffix = c("_transaltor", "_author"))

# Create clean author/illustrator/translator fields
stories_master <- stories_master %>%
                    mutate(translator_name = ifelse(derivation_type_updated == "original", "" , author_name_transaltor))

stories_master$author_name_updated <- str_replace(stories_master$author_name_author, " NA", "")
stories_master$illustrator_name_updated <- str_replace(stories_master$illustrator_name, " NA", "")
stories_master$translator_name_updated <- str_replace(stories_master$translator_name, " NA", "")

# Publisher story indicator
stories_master$publisher_type <- ifelse(is.na(stories_master$publisher), "UGC", "Publisher")

# Normalize data fields for use
stories_master$reading_level_updated <- as.factor(stories_master$reading_level_updated)
stories_master$language_name <- as.factor(stories_master$language_name)
stories_master$derivation_type_updated <- as.factor(stories_master$derivation_type_updated)
stories_master$publisher <- as.factor(stories_master$publisher)
stories_master$publisher_type <- as.factor(stories_master$publisher_type)

stories_final <- stories_master %>%
                    select(story_id, title, english_title, reading_level_updated, language_name,
                           synopsis, ancestry_updated, derivation_type_updated, reads, 
                           publisher, publisher_type, category_name, tag_name, is_recommended_story,
                           author_name_updated, illustrator_name_updated, translator_name_updated)
```

##### Summary of the dataset

```{r stories summary}
# Summary of the dataset
glimpse(stories_final)

# No. of stories
length(unique(stories_final$story_id))

# No. of Langugages
length(unique(stories_final$language_name))

# All reading levels
unique(stories_final$reading_level_updated)

# All publisher names
unique(stories_final$publisher_type)

# All story types
unique(stories_final$derivation_type_updated)

# No. of original stories
length(unique(stories_final$ancestry_updated))

# Total Reads
sum(stories_final$reads)

# Avg. Reads
mean(stories_final$reads)

# Preview of the dataset
head(stories_final)
```

##### Exploratory plots

We look at different cuts of the data to understand the compostion and device an approach.

```{r plots 1}
# Languages with most stories
top_language <- stories_final %>%
                    group_by(language_name) %>%
                    summarise(count = n()) %>%
                    mutate(percent = prop.table(count) * 100) %>%
                    arrange(desc(count)) %>%
                    head(10)

# Create the plot
ggplot(top_language, aes(x = reorder(language_name, -count), y = percent, fill = language_name)) +
    geom_bar(stat = "identity", width = 0.5) +
    geom_text(aes(y = percent + 1, label = paste0(sprintf(percent, fmt = "%.2f"), '%')), 
              position = position_dodge(width = .9), size = 3) +
    theme(plot.title = element_text(hjust = 0.5),
          axis.text.x = element_text(angle = 50, hjust = 1)) +
    labs(title = "Most publsished languges", x = "Language", y = "% Stories") +
    guides(fill = FALSE)


# Most read languages
top_language <- stories_final %>%
                    group_by(language_name) %>%
                    summarise(reads = sum(reads)) %>%
                    mutate(percent = prop.table(reads) * 100) %>%
                    arrange(desc(reads)) %>%
                    head(10)

# Create the plot
ggplot(top_language, aes(x = reorder(language_name, -reads), y = percent, fill = language_name)) +
    geom_bar(stat = "identity", width = 0.5) +
    geom_text(aes(y = percent + 1.8, label = paste0(sprintf(percent, fmt = "%.2f"), '%')), 
              position = position_dodge(width = .9), size = 3) +
    theme(plot.title = element_text(hjust = 0.5),
          axis.text.x = element_text(angle = 50, hjust = 1)) +
    labs(title = "Top read languages", x = "Language", y = "% Reads") +
    guides(fill = FALSE)

# 24% of all books published are in english
# 60% of all stories read are in the english
```

```{r plots 2}
# Count and Readership by story levels
top_level <- stories_final %>%
                    group_by(reading_level_updated) %>%
                    summarise(count = n(), reads = sum(reads)) %>%
                    mutate(percent_count = prop.table(count), percent_reads = prop.table(reads))

# Format uinque fields
top_level$count <- accounting(top_level$count, format = "d")
top_level$reads <- accounting(top_level$reads, format = "d")
top_level$percent_count <- percent(top_level$percent_count)
top_level$percent_reads <- percent(top_level$percent_reads)

# Create table
formattable(top_level, list(count = color_bar("orange"),
                            reads = color_bar("lightgreen"),
                            percent_count = color_tile("transparent", "lightblue"),
                            percent_reads = color_tile("transparent", "pink")))

# 47% of L1 stories have 38% readership
```

```{r plots 3}
# Count and Readership by recommendation of story
recco <- stories_final %>%
                group_by(is_recommended_story) %>%
                summarise(count = n(), reads = sum(reads)) %>%
                mutate(percent_count = prop.table(count), percent_reads = prop.table(reads))

# Format uinque fields
recco$count <- accounting(recco$count, format = "d")
recco$reads <- accounting(recco$reads, format = "d")
recco$percent_count <- percent(recco$percent_count)
recco$percent_reads <- percent(recco$percent_reads)

# Create table
formattable(recco, list(count = color_bar("orange"),
                        reads = color_bar("lightgreen"),
                        percent_count = color_tile("transparent", "lightblue"),
                        percent_reads = color_tile("transparent", "pink")))

# All publisher stories and only publisher stories are recommended stories
# 32% of the stories are recommended with have 66% reads
```

```{r plots 4}
# Count and Readership by original or derivative
orig <- stories_final %>%
            group_by(derivation_type_updated) %>%
            summarise(count = n(), reads = sum(reads)) %>%
            mutate(percent_count = prop.table(count), percent_reads = prop.table(reads))

# Format uinque fields
orig$count <- accounting(orig$count, format = "d")
orig$reads <- accounting(orig$reads, format = "d")
orig$percent_count <- percent(orig$percent_count)
orig$percent_reads <- percent(orig$percent_reads)

# Create table
formattable(orig, list(count = color_bar("orange"),
                        reads = color_bar("lightgreen"),
                        percent_count = color_tile("transparent", "lightblue"),
                        percent_reads = color_tile("transparent", "pink")))

# 28% of the stories are originals which have 60% reads
```

```{r plots 5}
# Look at readership and count by organisations
org_name <- stories_final %>%
                group_by(publisher) %>%
                filter(publisher != "NA") %>%
                summarise(count = n(), reads = sum(reads)) %>%
                mutate(percent_count = prop.table(count), percent_reads = prop.table(reads)) %>%
                arrange(desc(reads))

# Format uinque fields
org_name$count <- accounting(org_name$count, format = "d")
org_name$reads <- accounting(org_name$reads, format = "d")
org_name$percent_count <- percent(org_name$percent_count)
org_name$percent_reads <- percent(org_name$percent_reads)

# Create table
formattable(org_name, list(count = color_bar("orange"),
                        reads = color_bar("lightgreen"),
                        percent_count = color_tile("transparent", "lightblue"),
                        percent_reads = color_tile("transparent", "pink")))

# 87% of all publisher stories on the platform are by Pratham Books
# 88% of all publisher stories reads on the platform are for Pratham Books
```

#### Exploratory Data Analysis - Categories

##### Summary of the dataset

```{r category summary}
# Summary of the dataset
glimpse(categories)

# No. of unique categories
length(unique(categories$category_name))

# List of unique categories
unique(categories$category_name)
```

##### Exploratory plots

```{r plots 6}
# Most used categories
top_categories <- categories %>%
                    group_by(category_name) %>%
                    summarise(count = n()) %>%
                    mutate(percent = prop.table(count) * 100) %>%
                    arrange(desc(count)) %>%
                    head(10)

# Create the plot
ggplot(top_categories, aes(x = reorder(category_name, -count), y = percent, fill = category_name)) +
    geom_bar(stat = "identity", width = 0.5) +
    geom_text(aes(y = percent + 1, label = paste0(sprintf(percent, fmt = "%.2f"), '%')), 
              position = position_dodge(width = .9), size = 3) +
    theme(plot.title = element_text(hjust = 0.5),
          axis.text.x = element_text(angle = 50, hjust = 1)) +
    labs(title = "Top Categories", x = "Category", y = "%") +
    guides(fill = FALSE)
```

#### Exploratory Data Analysis - Tags

##### Summary of the dataset

```{r tag summary}
# Summary of the dataset
glimpse(tags)

# No. of unique tags
length(unique(tags$tag_name))
```

##### Exploratory plots

```{r plots 7}
# Most used categories
top_tags <- tags %>%
                group_by(tag_name) %>%
                summarise(count = n()) %>%
                mutate(percent = prop.table(count) * 100) %>%
                arrange(desc(count)) %>%
                head(10)

# Create the plot
ggplot(top_tags, aes(x = reorder(tag_name, -count), y = percent, fill = tag_name)) +
    geom_bar(stat = "identity", width = 0.5) +
    geom_text(aes(y = percent + .1, label = paste0(sprintf(percent, fmt = "%.2f"), '%')), 
              position = position_dodge(width = .9), size = 3) +
    theme(plot.title = element_text(hjust = 0.5),
          axis.text.x = element_text(angle = 50, hjust = 1)) +
    labs(title = "Top Tags", x = "Tag", y = "%") +
    guides(fill = FALSE)
```

#### Exploratory Data Analysis - Stories + Pages + Illustrations

##### Cleaning the dataset

```{r stories_pages_illustrations_master}
# stories_pages_illustrations_master dataset
# Clean-up and transform dataset for analysis
stories_pages_illustrations_master <- 
            stories_pages_illustrations_master %>%
                mutate(reading_level_updated = paste0("L", reading_level + 1),
                       # Normalize the reading_level field
                       ancestry_updated = ifelse(is.na(ancestry), story_id, ancestry),
                       # Normalize the ancestry field
                       derivation_type_updated = ifelse(is.na(derivation_type), "original", derivation_type),
                       # Get recommended story indicator
                       is_recommended_story = ifelse(is.na(publisher), F, T)
                      )

# Clean ancestry for double parent stories
stories_pages_illustrations_master <- 
        stories_pages_illustrations_master %>% 
                separate(ancestry_updated, c("ancestry_updated"), "/", extra = "drop", fill = "right")

stories_pages_illustrations_master$ancestry_updated <- as.numeric(stories_pages_illustrations_master$ancestry_updated)

# Processing data to creater master file for dashboard
# Join all the relevant data files
spi_master <- stories_pages_illustrations_master %>%
                    left_join(categories_grouped, by = "story_id") %>%
                    left_join(tags_grouped, by = "story_id") %>%
                    left_join(authors_grouped, by = "story_id") %>%
                    left_join(illustrators_grouped, by = "story_id") %>%
                    left_join(authors_grouped, by = c("ancestry_updated" = "story_id"), suffix = c("_transaltor", "_author"))

# Create clean author/illustrator/translator fields
spi_master <- spi_master %>%
                    mutate(translator_name = ifelse(derivation_type_updated == "original", "" , author_name_transaltor))

spi_master$author_name_updated <- str_replace(spi_master$author_name_author, " NA", "")
spi_master$illustrator_name_updated <- str_replace(spi_master$illustrator_name, " NA", "")
spi_master$translator_name_updated <- str_replace(spi_master$translator_name, " NA", "")

# Publisher story indicator
spi_master$publisher_type <- ifelse(is.na(spi_master$publisher), "UGC", "Publisher")

# Normalize data fields for use
spi_master$reading_level_updated <- as.factor(spi_master$reading_level_updated)
spi_master$language_name <- as.factor(spi_master$language_name)
spi_master$derivation_type_updated <- as.factor(spi_master$derivation_type_updated)
spi_master$publisher <- as.factor(spi_master$publisher)
spi_master$publisher_type <- as.factor(spi_master$publisher_type)

# Clean the Story Page text
cleanHTML <- function(htmlString) {
  return(gsub("<.*?>", "", htmlString))
}

spi_master$story_page_updated <- cleanHTML(spi_master$story_page)

# Create Master File for usage
spi_final <- spi_master %>%
                    select(story_id, title, english_title, reading_level_updated, language_name,
                           synopsis, ancestry_updated, derivation_type_updated, reads, 
                           publisher, publisher_type, category_name, tag_name, is_recommended_story,
                           author_name_updated, illustrator_name_updated, translator_name_updated,
                           story_page_position, story_page_updated, story_page_type, illustration_name)

glimpse(spi_final)
```

##### Summary of the dataset

```{r, warning = FALSE}
# No. of pages/story
length(spi_final$story_page_position[spi_final$story_page_type == "StoryPage"])/length(unique(spi_final$story_id))

# No. of illustations/story
length(spi_final$illustration_name[!is.na(spi_final$illustration_name)])/length(unique(spi_final$story_id))

# Max pages in a story
max(spi_final$story_page_position)

# No. of unique illustrations
length(unique(spi_final$illustration_name))
```

```{r}
# Store the cleaned file for pre-processing
write_csv(spi_final, "../data/processed/spi_cleaned.csv")
```
